% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{array}
\usepackage{vampire_induction}
\usepackage{graphicx}
\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{xargs}
\usepackage{adjustbox}
\usepackage{proof}
\usepackage{cancel}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{mdframed}
\usepackage{hhline}

{
\small 
\newcolumntype{R}[2]{%
	>{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
	l%
	<{\egroup}%
}
\newcommand*\rot[1]{\multicolumn{1}{R{#1}{1em}}}


\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
 %                      {8\p@ \@plus 4\p@ \@minus 4\p@}%
 %                      {\normalfont\large\bfseries\boldmath
 %                       \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
\makeatother
\newcommand{\AV}[1]{{\color{red}{AV: #1}}}

\newcommand{\todo}[1]{\emph{\color{red}{TODO: #1}}}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Structural Induction in Superposition Reasoning}
%
\titlerunning{ }
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{M\'arton Hajd\'{u}\inst{1} \and Petra Hozzov\'a\inst{1} \and
  Laura Kov\'acs\inst{1\and 2} \and Andrei Voronkov\inst{3,}\inst{4}}
%
\authorrunning{ }
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  TU Wien, Austria \and  Chalmers University of Technology, Sweden
  \and
  University of Manchester, UK
  \and
  EasyChair
}

%

\maketitle              % typeset the header of the contribution
%
%\vspace*{-1.3em}
\begin{abstract}
%\keywords{First keyword  \and Second keyword \and Another keyword.}

Formally ensuring software reliability is challenging due to the growing complexity of software in terms of its size, used unbounded and inductive data types and supported workflow. As manual code analysis is not a viable solution, automated reasoning approaches are needed for proving software correct. There are several prominent methods addressing this challenge by (i) translating parts of a system to a logic-based formalism and (ii) using automated theorem proving to validate certain safety and liveness program properties.

In this context a vital ingredient in automated reasoning is inductive reasoning, which can/must be used whenever repetition is present in a software component, e.g. loops or recursive data structures in computer programs or feedback loops in electric circuits. The two main branches in the literature
are explicit and implicit induction. In this thesis focus on the former approach which is based broadly on the Noetherian well-founded induction principle. There is a plethora of research already done in this area with several automated theorem provers as well as proof assistants are equipped with inductive reasoning techniques. Some of the earliest ones are \textsc{Nqthm/ACL2}, \textsc{Inka} and \textsc{Oyster/CLAM}. Recently, modern automated theorem provers, such as \textsc{ZipperPosition}, \textsc{CVC4} and \textsc{Vampire}, started to incorporate inductive types and reasoning.

The aim of the proposed thesis is to advance the state-of-the art in inductive reasoning within first-order theorem proving, by using automating reasoning with recursive function definitions over inductive data types.
\end{abstract}

\section{Introduction}
\input{intro.tex}

\section{Motivating Examples}
\label{sec:motivating}
\input{motivating_example.tex}

\section{Preliminaries}
\input{preliminaries.tex}

\section{Function definitions and Structural Induction}
\label{sec:fn_defs}
\input{function_defs.tex}

\section{Function definition rewriting in Superposition}
\label{sec:fn_def_rewriting}
\input{function_def_rewriting.tex}

\section{Induction in Saturation-based proof search}

\section{Experiments}
\label{sec:experiments}
\input{experiments.tex}

\section{Related Work}
\input{related_work.tex}

\section{Conclusions}

\section*{Acknowledgements}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%

\bibliographystyle{splncs04}
\bibliography{bib}
 
\end{document}

